---
title: "Class 5: Data visualization with ggplot"
author: "Siyu Xie (PID:16438448)"
format: pdf
toc: true
---

Today we will have our first play with the **ggplot2** package - one of the most popular graphics packages on the planet. :D

There are many plotting systems in R. These include so-called *"base"* plotting/graphics.

```{r}
plot(cars, xlab="speed(MPH)", ylab="stopping distance(ft)")
```

Base plot is generally rather short code and somewhat dull plots - but it is always there for you and is fast for big datasets.

If I want to use **ggplot2** it takes some more work.

```{r}
#ggplot(cars)
```

I need to install the package first to my computer. To do this I can use the function `install.packages("ggplot2")`

Every time I want to use a package I need to load it up with a `library()` call.

```{r}
library (ggplot2)
```

Now finally I can use ggplot!

```{r}
ggplot(cars)
```

Every ggplot has at least 3 things:

-   **data** (the data.frame with the data you want to plot)
-   **aes** (the aesthetic mapping of the data to the plot)
-   **geom** (how do you want the plot to look, points, lines, etc.)

```{r}
head(cars)
```
```{r}
bp <- ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point()
bp
```

```{r}
ggplot(cars) +
  aes(x=speed, y=dist) +
  geom_point() +
  geom_smooth()
```

I want a linear model and no standard error bounds/shaded standard error region shown on my plot. I also want nicer axis labels, a title, etc.


```{r}
bp + geom_smooth(se=FALSE, method="lm") +
  labs(title="Speed and Stopping Distances of Cars",
      x="Speed (MPH)",
      y="Stopping Distance (ft)",
      caption = "From the 'cars' dataset")+
  theme_bw()
```

## A more complicated scatter plot

Here we make a plot of gene expression data

```{r}
url <- "https://bioboot.github.io/bimm143_S20/class-material/up_down_expression.txt"
genes <- read.delim(url)
head(genes)
```

> Q.Use the nrow() function to find out how many genes are in this dataset. What is your answer?

```{r}
nrow(genes)
```

> Q. Use the colnames() function and the ncol() function on the genes data frame to find out what the column names are (we will need these later) and how many columns there are. How many columns did you find?

```{r}
colnames(genes)
ncol(genes)
```

> Use the table() function on the State column of this data.frame to find out how many ‘up’ regulated genes there are. What is your answer?

```{r}
table(genes$State)
```
```{r}
sum(genes$State =="up")
```

> Q.Using your values above and 2 significant figures. What fraction of total genes is up-regulated in this dataset?

```{r}
round(sum(genes$State =="up")/nrow(genes) * 100, 2)
```

```{r}
n.gene <- nrow(genes)
n.up <- sum(genes$State =="up")

up.percent <- n.up/n.gene * 100
round (up.percent, 2)
```

```{r}
head(genes, 2)
```

```{r}
p <- ggplot(genes) +
  aes(x=Condition1, y=Condition2, col=State) +
  geom_point()
p
```

```{r}
p + labs(title="Gene Expression Changes Upon Drug Treatment",
  x="Control (no drug)",
  y="Treated (drug)") +
  scale_color_manual( values=c("pink","grey","turquoise"))

```
## Explore the gapminder dataset

Here we will load up the gapminder dataset to get practice with different aes mappings.

```{r}
url <- "https://raw.githubusercontent.com/jennybc/gapminder/master/inst/extdata/gapminder.tsv"

gapminder <- read.delim(url)
```

>Q. how many rows and columns can be found in this dataframe?

```{r}
row <- nrow(gapminder)
column <- ncol(gapminder)

row
column
```

```{r}
head(gapminder)
```
>Q. How many continents are there in the dataset?

```{r}
table(gapminder$continent)
```


I could use the `unique()` function...
```{r}
length(unique(gapminder$continent))
```


>Q. How many countries are there in this dataset?

```{r}
length(unique(gapminder$country))
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(alpha=0.2)
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point(col="turquoise", alpha=0.3)
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=continent, size= pop) +
  geom_point(alpha=0.5)
```
```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp, col=pop) +
  geom_point(alpha=0.8)
```

```{r}
#install.packages("dplyr")
```

```{r}
library(dplyr)
gapminder_2007 <- filter(gapminder, year==2007)
head(gapminder_2007)
```

plot of 2007 with population and continent data
```{r}
ggplot(gapminder_2007) + 
  aes(x=gdpPercap, y=lifeExp, col=continent, size= pop) +
  geom_point(alpha=0.5)
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point() +
  facet_wrap(~year)
```

```{r}
ggplot(gapminder) +
  aes(x=gdpPercap, y=lifeExp) +
  geom_point() +
  facet_wrap(~continent)
```

