---
title: "class06HW_function"
author: Siyu Xie (PID:A16438448)
format: pdf
toc: true
---

## Analysis code snippet

```{r}
#install.packages("bio3d")
```

> Can you improve this analysis code?

```{r}
library(bio3d)
s1 <- read.pdb("4AKE")  # kinase with drug
s2 <- read.pdb("1AKE")  # kinase no drug
s3 <- read.pdb("1E4Y")  # kinase with drug

s1.chainA <- trim.pdb(s1, chain="A", elety="CA")
s2.chainA <- trim.pdb(s2, chain="A", elety="CA")
s3.chainA <- trim.pdb(s1, chain="A", elety="CA")
s1.b <- s1.chainA$atom$b
s2.b <- s2.chainA$atom$b
s3.b <- s3.chainA$atom$b
plotb3(s1.b, sse=s1.chainA, typ="l", ylab="Bfactor")
plotb3(s2.b, sse=s2.chainA, typ="l", ylab="Bfactor")
plotb3(s3.b, sse=s3.chainA, typ="l", ylab="Bfactor")


```

## Improved R function

**Documentation**:

**1. comments on the inputs to the function**

PDB_data: the PDB file name extracted from the PDB data for analysis
chain_name: chain interested in analysis

**2. what the function does and how to use it**

The function analyzes protein drug interactions by reading any protein's PDB data. To use the function, input the protein's PDB data name and the corresponding chain ID.

**3. what's the output of the function**

The function outputs a plot displaying the B-factor trends of the specified protein upon drug interactions.

```{r}
protein <- function(PDB_data, chain_name){
  library(bio3d)
# reading PDB file and trim the data
    protein_chain <- trim.pdb(read.pdb(PDB_data), chain = chain_name, elety ="CA")
# getting b factors from data
    b_factor <- protein_chain$atom$b
# plotting b factor trend for the specified protein
    plotb3(b_factor, sse=chain_name, typ="l", ylab="Bfactor")
}

protein("4AKE", "A")
    
```




