---
title: "class06: R function"
author: "Siyu Xie (A16438448)"
format: pdf
toc: true
---

All functions in R have at least 3 things:
- a **name** (you get to pick this)
- input **argument** (there can be only one or loads - again your call)
- the **body** (where the work gets done, this code between the curly brackets {})

## A first silly function

lets write a function to add some numbers. We can call it`add()`

```{r}
add <- function (x) {
  y <- 10
  x+y
  
}  

```

Can I just use my new function?

```{r}
add(1)
```

Let's make it a bit more flexible.

```{r}
add <- function (x,y=1) {
  x+y
}  

add(x=10, y=10)

```

```{r}
add(10)
```

## 2nd example grade() function

Write a function to grade student work.

We will start with a simple version of the problem and the following example student vectors:

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

Start with student1

```{r}
mean(student2, na.rm=TRUE)
```

```{r}
mean(student3, na.rm=TRUE)
```

```{r}
student1
```

google told me about min() and max().

```{r}
min(student1)
```
```{r}
which.min(student1)
```
Indicate the location/index where the min occurs

```{r}
student1[which.min(student1)]
```
```{r}
student1[-8]
```

our first working snippet that drops the lowest score and calculates the mean
```{r}
mean(student1[-which.min(student1)])
```

```{r}
x <- student2
mean(x[-which.min(x)], na.rm=T)
```
```{r}
x <- student3
mean(x[-which.min(x)], na.rm=T)
```

Our approach to the NA problem(missing homeworks): we can replace all NA values with zero.

1st task is find the NA values (i.e. where are they in the vector)
```{r}
x <- student2
x
is.na(x)
```

I have found the NA(TRUE) values from `is.na()` now I want to make them equal to zero. (mask them)

```{r}
x
x[is.na(x)] <- 0
```

```{r}
y <- 1:5
y
y[y>3] <-0
y
```
I want to combine the `is.na(x)` with making these elements equal to zero. And then take this "masked" (vector of student scores with NA values as zero) and drop the lowest and get the mean.

```{r}
x <- student3
x[is.na(x)] <- 0
mean(x[-which.min(x)])
```
Now I can turn my most awesome snippet into my first function

```{r}
grade <- function(x) {
  #Make NA (missing work) equal to 0
  x[is.na(x)] <- 0
  #Drop lowest score and get the mean
  mean(x[-which.min(x)])
}
```

```{r}
grade(student1)
```
#Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput” [3pts]

```{r}
url <- "https://tinyurl.com/gradeinput"
  gradebook <- read.csv(url,row.names = 1)
  head(gradebook)
```
The `apply()` function in R is super useful but can be a little confusing to begin with. Lets have a look how it works.

```{r}
ans <- apply(gradebook, 1, grade)
ans
```
#Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook?

```{r}
  which.max(ans)
  max(ans)
```
#Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall?

```{r}
lowestscores <- apply(gradebook, 2, mean, na.rm=T)
lowestscores
which.min(lowestscores)
min(lowestscores)

```

#Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)?

```{r}
#ans
cor(gradebook$hw1,ans)
```
#mask gradebook to assign NA to 0 scores
```{r}
mask <- gradebook
mask[is.na(mask)] <- 0
mask
```

```{r}
#ans
cor(mask$hw2,ans)
```

```{r}
highestcor <- apply(mask, 2, cor, y=ans)
highestcor
which.max(highestcor)
```

